---

---
Machine Learning - Peer Graded Assignment 
Yilma Adem
Date: July 15, 2017
----
## Summary
The goal of this project is to predict the model using pml training and test dataset. The dataset is sourced from  'Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements' by 6 researchers. Based on this source  a dataset is  with 5 classes (sitting-down, standing-up, standing, walking, and sitting) collected on 8 hours of activities of 4 healthy subjects. The dataset collected from devices such as Jawbone Up, Nike FuelBand, and Fitbit which relatively inexpensive. Complete history about HAR and the datasets can found http://groupware.les.inf.puc-rio.br/har.

Decision Tree and Randomforest have chosen to do model fitting and predictive analysis. However, based on accurancy rate, Randomforest  model is found to be good. The cons for randomforest is speed and amount memory needs.

## Packages
```{r echo= TRUE}
suppressMessages(library(caret))
suppressMessages(library(ggplot2))
suppressMessages(library(randomForest))
suppressMessages(library(rpart))
suppressMessages(library(Hmisc))
set.seed(12345)
```

### Data Loading , preparating and Exploring

```{r echo=TRUE}
dataTraining <-read.csv("pml-training.csv", na.strings = "")
dataTesting <- read.csv("pml-testing.csv", na.strings = "NA")
# creating a partition with the training dataset
inTrain <- createDataPartition(dataTraining$classe, p=0.7, list=FALSE)
training <- dataTraining[inTrain,]
testing <- dataTesting[-inTrain,]
dim(training);dim(testing)

## The dataset has  observation and 160 variables. There are substantial number of NAs near zero variables. So, removing these variables benifical to the model fiting and predictions.

Zeros <- nearZeroVar(training)
training <- training[, -Zeros]
testing <- testing[, -Zeros]
dim(training);dim(testing)

#removing NAs
NAs <-sapply(training, function(x) mean(is.na(x))) > 0.946
training <- training[, NAs==FALSE]
testing <- testing[,  NAs==FALSE]

#remove variables in columns 1 to 5
training <- training[, -(1:5)]
testing<- testing[, -(1:5)]
dim(training);dim(testing)
# Correlation among variables using plot.
qplot(total_accel_dumbbell, gyros_belt_y,  colour = classe, data=training)
```

###Model Fitting and Predicting using Decission Tree 

```{r echo=TRUE}
# Model fitting with Decission Tree
modFitDT <- rpart(classe ~.,data=training, method="class")

# Predicting 
PredictDT  <- predict(modFitDT, newdata=testing)
PredictDT
# Decision Tree accuracy is 0.74
```
### Model fitting and predicting  with Randomforest 
```{r echo=TRUE}
#To make parallel processing since randomforest using up memory and time
suppressMessages(library(doParallel))
#Find out how many cores are available
 cores<-detectCores()
 cl <- makeCluster(cores[1]-1)
#Register cluster
registerDoParallel(cl)
# Model Fitting using Random Forest
 set.seed(12345)
 conrf <- trainControl(method = "CV", number = 3, verboseIter = FALSE)
 modFitrf <- train(classe ~ ., method = "rf", data = training, trControl = conrf)
 modFitrf$finalModel
 
 # predicting using Random Forest with 8 observations of testing dataset
 Predrf <- predict(modFitrf, newdata = testing)
 Predrf
 # Predication accuracy  
  ## Randomforest accuracy is 0.99

```
###Conclusion
 With  accuracy 0.99 , Random forest regression model will be applied to predict the testing dataset.